\chapter*{Preface}

This thesis report marks the end of this project, on which I have worked for a year. The year started at the IBM Austin Research Lab, after Peter Hofstee invited me to work with him on emerging coherently attached FPGA accelerators. During my six months in Austin, Peter was always ready to help and chat. Also in the period after Austin was Peter always available. Peter, it has been a very pleasant journey and I sincerely hope our paths cross again. I can safely say that you not only have been a tremendous supervisor, but also a good friend.

Because the project is in the field of FPGAs, I had many interesting discussions with Andrew Martin. Andrew is a research staff member and developed a ready-valid design methodology for FPGA design. Andy, I would like to thank you for your support during the design and implementation phase. Your methodology is the missing link for FPGA design.

Dorus, Eric, Jeremy, and Jinho, the three o'clock stretch session must live on, but in different countries and time zones. I enjoyed my time in Austin very much and that is thanks to you.

I would like to thank the Universiteitsfonds for providing funding for the six months I spent in Austin. Without this support, it would have been difficult to have had this experience.

I would like to thank my parents for their life-long support and faith in me.

Finally, I would like to thank Fjóla for always being there for me when I needed her the most and making me a better person every day.
